Generalized Universe Hierarchies and First-Class Universe Levels

### 2021-02-17 paper.tex
progress

\theoremstyle{definition}
\newtheorem{mydefinition}{Definition}
\newtheorem{myexample}{Example}

@@

\newcommand{\Set}[1]{\mathsf{Set_{#1}}}
\newcommand{\Seti}{\mathsf{Set}}
\newcommand{\refl}{\mathsf{refl}}
\newcommand{\Con}{\mathsf{Con}}
\newcommand{\Ty}{\mathsf{Ty}}
\newcommand{\Tm}{\mathsf{Tm}}
\newcommand{\Sub}{\mathsf{Sub}}
\newcommand{\emptycon}{\scaleobj{.75}\bullet}
\newcommand{\U}{\mathsf{U}}
\newcommand{\El}{\mathsf{El}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\ext}{\triangleright}
\newcommand{\blank}{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}}
\newcommand{\mi}[1]{\mathit{#1}}
\newcommand{\p}{\mathsf{p}}
\newcommand{\q}{\mathsf{q}}

\renewcommand{\tt}{\mathsf{tt}}

@@

\begin{itemize}
\item Agda recently added a limited version cumulativity as an optional feature
  for universes, and the upcoming 2.6.2 version will extend the $\omega+1$
  universe hierarchy to $\omega*2$.
\item Coq added support for cumulative inductive types \cite{timany18cumulative}, and added a form of
  bounded universe polymorphism \cite{ziliani15unification}.
\end{itemize}

@@

\begin{enumerate}
\item In Section \ref{sec:ttgu} we describe models of type theories where
  universe levels may come from any set with a well-founded transitive ordering
  relation. We specify models as categories equipped with level-indexed diagrams
  of families, as a variation on categories with families. Each morphism of
  levels is mapped to a lifting operation on terms and types. By varying the
  preservation properties of lifting operations, we can describe a range of
  stratification features, from two-level type theory to cumulative universes.
\item We use induction-recursion to model the mentioned theories. Here, we model
  the strongest formulations for lifting and universes, namely cumulative
  universes with Russell-style type decoding. The main challenge is combining
  transfinite hierarchies with strictly type-preserving lifting operations.
\item In Section \ref{sec:ttfl} we describe type theories with internal types
  for levels and level morphisms, and extend the previous inductive-recursive
  semantics to cover these as well. Here, we can additionally represent various
  universe polymorphism features and level computations. We illustrate how
  predicative universe features in Coq and Agda could be justified in our framework.
\end{enumerate}

@@

We provide an Agda formalization of the contents of the paper. The formalization
is not complete, as we skip the proofs involving an over-abundance of equality
coercions (which are more suited to informal reasoning, using equality
reflection), and instead focus on the key points.

@@

We work in a Martin-Löf type which has the following features.
\begin{itemize}
  \item Two universes named $\Set0$ and $\Set1$, where $\Set0$ supports
    inductive-recursive types (IR) as specified by Dybjer and Setzer
    \cite{dybjer99finite}. The $\Set1$ universe is not essential and we only use
    it as a convenience feature, in this paper and in the Agda formalization. We
    may omit the universe indices if they can be inferred or if we work over
    arbitrary indices.
  \item Function extensionality and uniqueness of identity proofs
    (UIP). Additionally, we assume equality reflection in this paper, thus
    working in extensional type theory, to avoid noise from equality transports.
  \item We write function types as $(x : A)\to B$ with $\lambda\,x.\,t$
    inhabitants, and $\Sigma$-types as $(x : A) \times B$, with pairing as
    $(t,\, u)$ and projections as $t.1$ and $t.2$. We have $\top$ as the unit
    type with inhabitant $\tt$, and $\bot$ as the empty type. Propositional identity
    is written as $t = u$ with $\refl : t = t$.
  \item We occasionally use $\{x : A\} \to B$ for an Agda-like notation for
    function types with implicit arguments, and omit implicit applications
    whenever they can be inferred.

\end{itemize}

@@

In this section, we first describe notions of models for type theories with
generalized universes, and discuss several variations of universes and lifting
operations. Then, we pick a concrete variant (the ``strongest'', in a sense)
and construct a model for it in the metatheory.

@@

\begin{mydefinition}[Category with family (CwF), c.f. \cite{Dybjer96internaltype}]\label{def:cwf}
A CwF consists of the following data:
\begin{itemize}
\item A category with a terminal object. We denote the set of objects as $\Con :
  \Seti$ and use capital Greek letters starting from $\Gamma$ to refer to
  objects. The set of morphisms is $\Sub : \Con \to \Con \to \Seti$, and we use
  $\sigma$, $\delta$ and so on to refer to morphisms. The terminal object is
  $\emptycon$ with unique morphism $\epsilon : \Sub\,\Gamma\,\emptycon$. In
  initial models (that is, syntaxes) of type theories, objects correspond to
  typing contexts, morphisms to parallel substitutions and the terminal object to
  the empty context; this informs the naming scheme that we use here.
\item A \emph{family structure}, containing $\Ty : \Con \to \Seti$ and $\Tm :
  (\Gamma : \Con) \to \Ty\,\Gamma \to \Seti$, where $\Ty$ is a presheaf over the
  category of contexts and $\Tm$ is a presheaf over the category of elements of
  $\Ty$. This means that both types ($\Ty$) and terms ($\Tm$) can be
  substituted, and substitution has functorial action. We use $A$, $B$, $C$ to
  refer to types and $t$, $u$, $v$ to refer to terms, and use $A[\sigma]$ and
  $t[\sigma]$ for substituting types and terms. Additionally, a family
  structure has \emph{context comprehension} which consists of an operation
  $\blank\ext\blank : (\Gamma : \Con) \to \Ty\,\Gamma \to \Con$ together with an
  isomorphism $\Sub\,\Gamma\,(\Delta\ext A) \simeq ((\sigma : \Sub\,\Gamma\,\Delta)
  \times \Tm\,\Gamma\,(A[\sigma]))$ which is natural in $\Gamma$.
\end{itemize}
\end{mydefinition}

@@

notions of substitution extension, weakening
and variables, internally to any CwF.

\subsection{Morphisms, Lifts and Inclusions of Families}
\label{sec:morphisms}

morphisms, lifts (2ltt), inclusions, strict inclusion (cumulative)

@@

Level structures, thin direct semicat, vs thin cat, direct semicat, direct cat

>>>>

\theoremstyle{definition}
\newtheorem{mydefinition}{Definition}
\newtheorem{myexample}{Example}
\newtheorem{mylemma}{Lemma}

@@

\newcommand{\Set}[1]{\mathsf{Set_{#1}}}
\newcommand{\Seti}{\mathsf{Set}}
\newcommand{\refl}{\mathsf{refl}}
\newcommand{\Con}{\mathsf{Con}}
\newcommand{\Ty}{\mathsf{Ty}}
\newcommand{\Tm}{\mathsf{Tm}}
\newcommand{\Sub}{\mathsf{Sub}}
\newcommand{\emptycon}{\scaleobj{.75}\bullet}
\newcommand{\U}{\mathsf{U}}
\newcommand{\El}{\mathsf{El}}
\newcommand{\id}{\mathsf{id}}
\newcommand{\ext}{\triangleright}
\newcommand{\blank}{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}}
\newcommand{\mi}[1]{\mathit{#1}}
\newcommand{\p}{\mathsf{p}}
\newcommand{\q}{\mathsf{q}}
\newcommand{\Id}{\mathsf{Id}}
\newcommand{\Nat}{\mathsf{Nat}}
\newcommand{\Bool}{\mathsf{Bool}}
\newcommand{\true}{\mathsf{true}}
\newcommand{\false}{\mathsf{false}}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\Lift}{\mathsf{Lift}}
\renewcommand{\tt}{\mathsf{tt}}
\newcommand{\Acc}{\mathsf{Acc}}
\newcommand{\acc}{\mathsf{acc}}
\newcommand{\Lvl}{\mathsf{Lvl}}

@@

\begin{itemize}
\item Agda recently added a limited version cumulativity as an optional feature
  for universes \cite{agdadocs}, and the upcoming 2.6.2 version will extend the $\omega+1$
  universe hierarchy to $\omega*2$.
\item Coq added support for cumulative inductive types \cite{timany18cumulative}, and added a form of
  bounded universe polymorphism \cite{ziliani15unification}.
\end{itemize}

@@

\begin{enumerate}
\item In Section \ref{sec:ttgu} we describe models of type theories where
  universe levels may come from any set with a well-founded transitive ordering
  relation. We specify models as categories equipped with level-indexed diagrams
  of families, as a variation on categories with families. Each morphism of
  levels is mapped to a lifting operation on terms and types. By varying the
  preservation properties of lifting operations, we can describe a range of
  stratification features, from two-level type theory to cumulative universes.
\item We use induction-recursion to model the mentioned theories. We model
  the strongest formulations for lifting and universes, namely cumulative
  universes with Russell-style type decoding. The main challenge is combining
  transfinite hierarchies with strictly type-preserving lifting operations.
\item In Section \ref{sec:ttfl} we describe type theories with internal types
  for levels and level morphisms, and extend the previous inductive-recursive
  semantics to cover these as well. Here, we can additionally represent various
  universe polymorphism features and level computations. We illustrate how
  predicative universe features in Coq and Agda could be justified in our framework.
\end{enumerate}

@@

We provide an Agda formalization of the contents of the paper. The formalization
is not complete, as we skip proofs involving an excessive number of equality
coercions (which are more suited to informal reasoning, using equality
reflection), and instead focus on the key points.

@@

We work in a Martin-Löf type theory which has the following features.
\begin{itemize}
  \item Two universes named $\Set0$ and $\Set1$, where $\Set0$ supports
    inductive-recursive types (IR) as specified by Dybjer and Setzer
    \cite{dybjer99finite}. The $\Set1$ universe is not essential and we only use
    it as a convenience feature, in this paper and in the Agda formalization. We
    may omit the universe indices if they can be inferred or if we work over
    arbitrary indices.
  \item Function extensionality and uniqueness of identity proofs
    (UIP). Additionally, we assume equality reflection in this paper, thus
    working in extensional type theory, to avoid noise from equality transports.
  \item We write function types as $(x : A)\to B$ with $\lambda\,x.\,t$
    inhabitants, and $\Sigma$-types as $(x : A) \times B$, with pairing as
    $(t,\, u)$ and projections as $t.1$ and $t.2$. We have $\top$ as the unit
    type with inhabitant $\tt$, and $\bot$ as the empty type. Propositional identity
    is written as $t = u$ with $\refl : t = t$.
  \item We occasionally use $\{x : A\} \to B$ for an Agda-like notation for
    function types with implicit arguments, and omit implicit applications
    whenever they can be inferred. We may omit implicit function types
    altogether if it is clear where certain variables are quantified.
\end{itemize}

@@

In this section, we first describe notions of models for type theories with
generalized universes, and discuss several variations of universes and lifting
operations. Then, we pick a concrete variant (the strongest, in a sense)
and construct a model for it in the metatheory.

@@

\begin{mydefinition}
A \emph{category with family} (cwf) \cite{Dybjer96internaltype} consists of the following data:
\begin{itemize}
\item A category with a terminal object. We denote the set of objects as $\Con :
  \Seti$ and use capital Greek letters starting from $\Gamma$ to refer to
  objects. The set of morphisms is $\Sub : \Con \to \Con \to \Seti$, and we use
  $\sigma$, $\delta$ and so on to refer to morphisms. The terminal object is
  $\emptycon$ with unique morphism $\epsilon : \Sub\,\Gamma\,\emptycon$. In
  initial models (that is, syntaxes) of type theories, objects correspond to
  typing contexts, morphisms to parallel substitutions and the terminal object to
  the empty context; this informs the naming scheme that we use here.
\item A \emph{family structure}, containing $\Ty : \Con \to \Seti$ and $\Tm :
  (\Gamma : \Con) \to \Ty\,\Gamma \to \Seti$, where $\Ty$ is a presheaf over the
  category of contexts and $\Tm$ is a presheaf over the category of elements of
  $\Ty$. This means that both types ($\Ty$) and terms ($\Tm$) can be
  substituted, and substitution has functorial action. We use $A$, $B$, $C$ to
  refer to types and $t$, $u$, $v$ to refer to terms, and use $A[\sigma]$ and
  $t[\sigma]$ for substituting types and terms. Additionally, a family structure
  has \emph{context comprehension} which consists of an context extension
  operation $\blank\ext\blank : (\Gamma : \Con) \to \Ty\,\Gamma \to \Con$
  together with an isomorphism $\Sub\,\Gamma\,(\Delta\ext A) \simeq ((\sigma :
  \Sub\,\Gamma\,\Delta) \times \Tm\,\Gamma\,(A[\sigma]))$ which is natural in
  $\Gamma$.
\end{itemize}
\end{mydefinition}

@@

There are other ways for presenting the basic categorical structure of models,
which are nonetheless equivalent to cwfs, including natural models \cite{awodey18natural}
and categories with attributes \cite{cartmellthesis}. We use the cwf presentation for its
immediately algebraic character and closeness to syntax, as an explicit
substitution calculus can be directly read off the specification.

\begin{notation}As De Bruijn indices are hard to read, we will mostly use
nameful notation for binders. For example, assuming $\Nat : \{\Gamma : \Con\}
\to \Ty\,\Gamma$ and $\Id : \{\Gamma : \Con\}(A : \Ty\,\Gamma) \to
\Tm\,\Gamma\,A \to \Tm\,\Gamma\,A \to \Ty\,\Gamma$, we may write $\emptycon \ext
(n : \Nat) \ext (p : \Id\,\Nat\,n\,n)$ for a typing context, instead of using
numbered variables or cwf combinators as in $\emptycon \ext \Nat \ext
\Id\,\Nat\,\q\,\q$.
\end{notation}

\begin{notation}
In the following, we will denote families by ($\Ty$,$\Tm$) pairs and overload context
extension $\blank\ext\blank$ for different families.
\end{notation}

A family structure may be closed under certain \emph{type formers}. For example,
we may close a family over function types by assuming $\Pi : (A : \Ty\,\Gamma)
\to \Ty\,(\Gamma\ext A) \to \Ty\,\Gamma$ together with abstraction, application,
$\beta\eta$-rules, and equations for the action of substitution on type and term
formers.

\subsection{Morphisms, Liftings and Inclusions of Families}
\label{sec:morphisms}

In the rest of the paper we make use of categories equipped with possibly
multiple family structures, which serves as basis for specifying universe
hierarchies. However, it is not very useful to simply have multiple copies of
family structures together with their type formers. In that case, every
constructor and eliminator of every type former stays in the same family, and
there is no interaction between families, the most we can do is to mix them
together in typing contexts. In this subsection we describe several ways of
crossing between families.

\begin{mydefinition}
A \emph{family morphism} $F$ between ($\Tm_0$, $\Ty_0$) and ($\Tm_1$, $\Ty_1$)
families consists of natural transformations mapping types to types and terms to
terms, which preserves context extensions up to context isomorphism, i.e.\ we
have that $(\Gamma \ext F\,A) \simeq (\Gamma \ext A)$, where $\simeq$ denotes
existence of an invertible context morphism.
\end{mydefinition}

Family morphisms are restrictions of so-called \emph{weak morphisms}
\cite{dependentrightadjoints} (or \emph{pseudomorphisms}
\cite{kaposi2019gluing}) of cwfs: a weak morphism which has the identity action
on the base category is exactly a family morphism. A plain family morphism is
still not too interesting, so we add more features.

\begin{mydefinition}
A \emph{lifting} is a family morphism which is bijective on terms. Note that
we can drop the $\ext$-preservation condition from the specification of
liftings, since it follows from the invertible action on terms.
\end{mydefinition}

\begin{notation}
We write $\Lift : \Ty_0\,\Gamma \to \Ty_1\,\Gamma$ for the action of some lifting
on types, $\up : \Tm_0\,\Gamma\,A \to \Tm_1\,\Gamma\,(\Lift\,A)$ for the action
on terms, and $\down$ for the inverse action on terms.
\end{notation}

\todo{Mention dependent right adjoints}

\paragraph{Two-level type theory}
Assume family structures ($\Ty_0$, $\Tm_0$) and ($\Ty_1$, $\Tm_1$), and a
lifting between them. This corresponds to a basic version of \emph{two-level
  type theory} \cite{twolevel}. This theory can be interpreted as having ($\Ty_1$,
$\Tm_1$) as a metaprogramming layer which can generate object-level programs in
the ($\Ty_0$, $\Tm_0$) layer.

Lifted types correspond to types of object-level expressions; for
example, $\Bool_0 : \Ty_0\,\Gamma$ is the object-level type of Booleans, while
$\Lift\,\Bool_0$ is the meta-level type of $\Bool_0$-expressions, and $\Bool_1 :
\Ty_1\,\Gamma$ is the type of meta-level Booleans. It is possible to compute a
$\Lift\,\Bool_0$ from a $\Bool_1$. Given $b : \Tm_1\,\Gamma\,\Bool_1$, we can
construct
$\down(\mathsf{if}\,b\,\mathsf{then}\up\!\true_0\,\mathsf{else}\up\!\false_0) :
\Tm_0\,\Gamma\,\Bool_0$.

But there is no way to compute a $\Bool_1$ from a $\Lift\,\Bool_0$, since
$\Lift\,\Bool_0$ has no elimination rule in $\Ty_1$. Conceptually, terms
of $\Lift\,\Bool_0$ are \emph{expressions}, so they are not necessarily $\true$ or
$\false$, they can also be variables or neutral expressions, so the usual
Boolean elimination is not justified for them. In general, the setup ensures that
we can eliminate from positive types in $\Ty_1$ to their counterparts in
$\Ty_0$, but not the other way around, while negative types are preserved by
$\Lift$ up to type isomorphism.

Remarkably, the simple rules of two-level type theory appear to model a form of
generative two-stage compilation with dependent types. Comparing this system to
e.g.\ BER MetaOCaml \cite{kiselyov14metaocaml}, we can relate $\Lift$ to $\mathsf{code}$, $\up$
to the quasi-quotation operation $<\!\blank\!>$, and $\down$ to escape
$\sim\!\blank$.

While the staging aspect could be subject of future research, we currently focus
on ``sizing'' hierarchies instead of staging hierarchies. This means that we
want to eliminate from any family to any other family which is ``connected'' by
a morphism.

\begin{mydefinition} A \emph{family inclusion} is a lifting which preserves all type
and term formers. This assumes that every type former which is contained in the source family,
is also contained in the target family.
\end{mydefinition}

\noindent Some examples for preservation equations for type and term formers:
\begin{alignat*}{3}
  & \Lift\,(\Pi\,(x : A) B)   && =\,\,\,\,&& \Pi\,(x : \Lift\,A)(\Lift\, (B[x \mapsto\,\down\!x]))\\
  & \up(\lambda\,(x : A).\,t) && =&& \lambda\,(x : \Lift\,A).\up(t[x\mapsto\down\!x])\\
  & \Lift\,\Bool_0            && =&& \Bool_1\\
  & \up\true_0                && =&& \true_1
\end{alignat*}
In general, we can skip specifying preservation for $\down$, since it follows from $\up$ preservation
equations.

Assume an inclusion from ($\Ty_0$, $\Tm_0$) to ($\Ty_1$, $\Tm_0$). Now, we can
eliminate from $\Bool_0$ to $\Bool_1$. If we have some $b :
\Tm_0\,\Gamma\,\Bool_0$, we also have $\up\!b :
\Tm_1\,\Gamma\,(\Lift_1\,\Bool_0)$, hence $\up\!b :
\Tm_1\,\Gamma\,\Bool_1$. Then, we can use $\Bool_1$ elimination, as in
$\mathsf{if}\,\up\!b\,\,\mathsf{then}\,\true_1\,\mathsf{else}\,\false_1 :
\Tm_1\,\Gamma\,\Bool_1$. The $\up$ computation ensures that the eliminator
computes appropriately on canonical terms: if $b$ is $\true_0$, then we get
$\up\!\true_0 = \true_1$ as the if-then-else scrutinee.

A family inclusion corresponds to a \emph{cumulative hierarchy} consisting of
two families: every type former of the smaller family is included in the larger
family, with the same elimination rules.

\begin{mydefinition}\label{def:strict_inclusion}
A \emph{strict family inclusion} between ($\Tm_0$, $\Ty_0$) and ($\Tm_1$,
$\Ty_1$) is a family inclusion for which the following equations hold:
\begin{alignat}{2}
  & (\Gamma \ext \Lift\,A) &&= (\Gamma \ext A)     \label{eq:cumcon}    \\
  & \Tm_1\,\Gamma\,(\Lift\,A) &&= \Tm_0\,\Gamma\,A  \label{eq:cumtm}     \\
  & \up\!t &&= t                                    \label{eq:cumlift}
\end{alignat}

%% \begin{enumerate}
%% \item $(\Gamma \ext \Lift\,A) = (\Gamma \ext A)$
%% \item $\Tm_1\,\Gamma\,(\Lift\,A) = \Tm_0\,\Gamma\,A$
%% \item $\up\!t = t$
%% \end{enumerate}
\end{mydefinition}

A strict inclusion corresponds to Sterling's \emph{algebraic cumulativity}
\cite{sterling2019algebraic}. The additional equations are a matter of convenience: they allow us
to omit term liftings in informal syntax\footnote{In a proof assistant, we would
  still have to explicitly transport along the strict inclusion
  equations.}. Most of the time we can also omit level annotations
on term formers. For example, we have $\true_0 : \Tm_0\,\Gamma\,\Bool_0$,
but also $\true_0 : \Tm_0\,\Gamma\,(\Lift\,\Bool_0)$, hence $\true_0 :
\Tm_0\,\Gamma\,\Bool_1$. Moreover, $\true_0$ is definitionally equal to
$\true_1$, since $\true_0 =\,\up\!\true_0 = \true_1$. Thus, using simply $\true$
is fine whenever the family is clear from context.

The definitional equality of $\true_0$ and $\true_1$ is important; without it
canonicity would fail, since $\true_0$, $\false_0$, $\true_1$ and $\false_1$
would be four definitionally distinct inhabitants of $\Bool_1$. See Luo
\cite{luo2012notes} on related issues with cumulativity. This means that it is
not sufficient to specify a strict inclusion just by equations \ref{eq:cumcon}
and \ref{eq:cumtm} in Definition \ref{def:strict_inclusion}, we need $\up$ together
with equation \ref{eq:cumlift} to identify term formers in different families. The
other direction $\down\!t = t$ is immediately derivable.

@@

We would like to describe a range of setups with multiple families and morphisms
between them. In this subsection we describe the indexing structures for such
family diagrams. First, some preliminary definitions.

\begin{mydefinition} The \emph{accessibility predicate} on relations is defined by the following
inductive rules:
\begin{alignat*}{3}
  & \Acc : \{A : \Seti\}(R : A \to A \to \Seti) \to A \to \Seti \\
  & \acc : \{a : A\} \to ((\mi{a'} : A) \to R\,\mi{a'}\,a \to \Acc\,R\,\mi{a'}) \to \Acc\,R\,a
\end{alignat*}
\end{mydefinition}

\noindent An inhabitant of $\Acc\,R\,a$ proves that starting from $a : A$ all
descending $R$-chains must be finite. This is ensured by the universal property
of the inductive definition, as all inductive types must be well-founded.

\begin{mylemma} Accessibility is a mere proposition, i.e.\ all inhabitants of $\Acc\,R\,a$ are propositionally equal; see
  \cite[Section 10.3]{hottbook}.
\end{mylemma}

\begin{mydefinition} A relation $R : A \to A \to \Seti$ is \emph{well-founded} if $(a : A) \to \Acc\,R\,a$.
\end{mydefinition}

\begin{mydefinition} A \emph{level structure} consists of the following components:
\begin{alignat*}{3}
  & \Lvl                  &&: \Set0 \\
  & \blank\!<\!\blank     &&: \Lvl \to \Lvl \to \Set0 \\
  & \mathsf{<\!prop }     &&: (p\,q : i < j) \to p = q \\
  & \blank\!\circ\!\blank &&: j < k \to i < j \to i < k \\
  & \mathsf{<\!wf}        &&: (i : \Lvl) \to \Acc\,<\,i
\end{alignat*}
\end{mydefinition}

\noindent We overload $\Lvl$ to refer to a given level structure and also its
underlying set. In short, a level structure is a set together with a transitive
well-founded relation.

\begin{mydefinition}
A \emph{family diagram} over $\Lvl$ maps each $i : \Lvl$ to a family structure
($\Ty_i$, $\Tm_i$), and each $p : i < j$ to a strict inclusion
($\Lift_{i}^{j}\,p$, $\up_{i}^{j}p$, $\down_{i}^{j}p$) between ($\Ty_i$,
$\Tm_i$) and ($\Ty_j$, $\Tm_j$). Moreover, the mapping is functorial, so
$\Lift_{i}^{k}\,(p\circ q)\,A = \Lift_{j}^{k}\,p\,(\Lift_{i}^{j}\,q\,A)$, and
similarly for $\up_{i}^{j}p$ and $\down_{i}^{j}p$.
\end{mydefinition}

\begin{notation}
Sometimes we omit some of the $i$, $j$, $p$ annotations from type and term
liftings, if they are clear from context.
\end{notation}

Our choice of level structures and diagrams is motivated by the following
reasons. First, we do not need identity morphisms in levels, because they would
be mapped to trivial liftings, which are not interesting in our setting. Second,
we do not need proof-relevant level morphisms, since any parallel pair of family
liftings gives rise to isomorphic types. Concretely, given $p : i < j$ and $q :
i < j$ such that $p \neq q$, we have $\Tm_j\,\Gamma\,(\Lift\,p\,A) \simeq
\Tm_i\,\Gamma\,A \simeq \Tm_j\,\Gamma\,(\Lift\,q\,A)$, and since $\Lift\,p\,A$
and $\Lift\,q\,A$ are in the same family, we can internally prove them
isomorphic using function types and identity types.

%% \begin{notation}
%% Given $p : i < j$ we may write $\Lift_{i}^{j}\,p : \Ty_i\,\Gamma \to
%% \Ty_j\,\Gamma$ for type lifting, and $\up_{i}^{j}p$ and $\down_{i}^{j}p$
%% accordingly for term lifting. Sometimes we omit $i$, $j$ and $p$ annotations,
%% when there is no ambiguity.
%% \end{notation}

%% \end{definition}
